\speculake: Hiding Malware in Speculative Execution

### 2018-04-29 paper/architecture.tex
Added pass on limitations of speculative exec. Added zbuff data

\subsection{Limits of Speculative Execution}
We performed several experiments to investigate how much computation can be done
speculatively. We performed these experiments on an Intel~(TODO: Jack fill in).
TODO: Jack describe experiment and results:

- Experiment w/ graph of number of instructions
- Limitation is in the ROB (cite other work that supports this)

>>>>

\begin{figure}[t]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/Speculative_capacity.png}
    \caption{}
    \label{fig:spec-capacity}
\end{figure}

\subsection{Limits of Speculative Execution}
We performed several experiments to investigate how much computation can be done
speculatively. We performed these experiments on an Intel Xeon-1270, as well as 
an intel i5-7200U to verify the results across processor generations.

Speculative branching is not a new feature in processors~\cite{branching-hist}, 
and it has been integrated in multiple generations of commodity processors today. 
Accordingly speculative instruction capacity has fairly well defined limits, in 
line with known features of processor designs. We have determined that features 
like reorder buffers (ROB), physical register file (PRF), and cache miss timing
determines a strong limit on the number of instructions that can be accomplished
speculatively.

\subsubsection{Reorder Buffer Capacity} \label{sssec:ROB}
To determine these limits we first measured the capacity of the reorder buffer
using the method outlined by~\cite{intel-rob-capacity}. By performing two cache loads
with a variable number or instructions between them and watching the amount of 
time taked to complete this loop we can find a definitive number of instructions
at which the second cache access is outside of the ROB. Figure~\ref{fig:spec-capacity}
shows this step function at ~220 instructions for both Sandy Lake and Kaby Lake 
generation processors.

\subsubsection{Cache Miss Duration}
When executing instructions speculatively we rely on a load from a cache line that
has recently been evicted. We have examined the number of cycles a
cache miss takes to return by artificially evicting an item from cache and timing 
a request for that specific item. The distribution of cycle counts can be seen in
Figure~\ref{fig:cache-miss}, along with the CDF. For a large majority of cases, 
an item evicted from cache will have 300 clock cycles and often much more.

\subsubsection{Speculative Instruction Capacity}
To verify that the ROB defines an upper limit for speculative instruction capacity
we instrumented the trigger and payload to test variable length target functions. 
In addition we tested whether instruction complexity or data dependencies impact
the maximum work that can be completed. We found that the processor can identify 
some operations that have no effect on output registers and allows them to use the 
entire size provided by the ROB, as determined in section~\ref{sssec:ROB}. This includes 
idomatic no-ops, which have no effect on the registers, for example the instruction 
"xchg \%rax \%rax" or similarly "xor \%rax, \%rax". However, if an instruction
has the potential to cause data dependencies, it must use an entry in the PRF
until it reaches the correct stage in the execution pipeline. In this case, the number 
of instructions is limited by the PRF entries available

Most notably instructions making use of the extended x86 registers are still
valid within speculative context. Specifically Intel's hardware accelerated AES-NI
encryption and decryption instructions, which make use of the 128 bit registers.
As shown in Figure~\ref{fig:spec-capacity}, speculative environments can complete
a significant number of AES rounds -- more than enough to decrypt a full block 
using AES-CBC. We investigate the use of AES instructions in the spculative environment
further in Sections~\ref{subsec:decryption}.

\medskip

When executing speculatively, the number of instrutions completed 
will almost never be limited by cache miss duration. Instead
ROB size and processor mechanisms for resolving 
data dependencies define an upper boundary.

% TODO: 
% - Highlist instruction signal variablity (add v mul v aesdec v nop)?
% - Limitation is in the ROB (cite other work that supports this)
% - MORE PROCESSORS for testing?
% - ARM?
